Classless Java

### 2016-03-24 OOPSLA16/sections/Implementation.tex
Some spacing tweking.
unify title/paragraph/enumerate capitalization: The rule is: - Capitalization All Section/Subsection Titles. - No capitalization for paragraph/enumerate titles. - No capitalization for the first word after a colon ":" (There's difference in this point between American and British English, I think we should stick to one).

\begin{figure}[t]
\centering
\includegraphics[width=3in]{pdfs/lombok3.png}
\caption{The flow chart of \mixin annotation processing.
}
\label{fig:lombok}
\end{figure}

@@

\paragraph{Direct modification of the AST:}
Lombok alters the generation process of the class files,
by directly modifying the AST. Neither the source code is modified nor
new Java files are generated. Moreover, and probably more importantly,
Lombok supports generation of code \emph{inside} a class/interface,
which conventional Java annotation processors do not support. For
example, the standard \texttt{javax.annotation} processor, which is part of the
Java platform, only allows generation of \emph{new code}, and the
new code has to be written in \emph{new files}. Modification and/or
reinterpretation of existing code is not supported.

@@

\begin{figure*}[t]
\centering
\includegraphics[width=6in]{pdfs/screenshot4.png}
\caption{Generated methods shown in the Outline window of Eclipse and auto-completion.}
\label{fig:screenshot}
\end{figure*}

@@

\paragraph{Clarity against obfuscation:}
Preprocessors bring great power, which can easily be misused producing
code particularly hard to understand. Thus code quality and maintainability are reduced.
Compilation agents start from Java syntax, but they can reinterpret it.
Preserving the syntax avoids syntactic conflicts, and allows many
tools to work transparently.

@@

\paragraph{Syntax and type errors}:
Some preprocessors (like the C one) can produce syntactically invalid code.
Lombok ensures only syntactically valid code is produced. %; however, type errors can appear.
Classless Java additionally guarantees that no type errors are introduced
in generated code and client code. We discuss these two guarantees in
more detail next:

@@

\item{\bf Self coherence}: The generated code itself is well-typed. That is,
  type errors are not present in code the user has not written (for
  example \texttt{of} methods in Figure~\ref{fig:screenshot}).
In our case it means that either \mixin{} produces (in a controlled way) an
understandable error, or the interface can be successfully annotated and the generated code is well-typed.

\item{\bf Client coherence}: All the client code (for example method calls)
  that is well-typed before code generation is also well-typed after the generation.
The annotation just adds more behaviour without removing any functionality.

@@

\paragraph{Heir coherence}: Another form of guarantee that could be
useful in AST rewriting is heir coherence. That is, interfaces
(and in general classes) inheriting the instrumented code are
well-typed if they were well-typed without the instrumentation.
In a strict sense, our rewriting \emph{does not} guarantee heir coherence.  The reason
is that this would forbid adding any (default or abstract) method to
the annotated interfaces, or even doing type refinement. Indeed consider
the following:

>>>>

\begin{figure}[t]
\saveSpaceFig
\centering
\includegraphics[width=3in]{pdfs/lombok3.png}
\caption{The flow chart of \mixin annotation processing.
}
\label{fig:lombok}
\saveSpaceFig
\end{figure}

@@

\paragraph{Direct modification of the AST}
Lombok alters the generation process of the class files,
by directly modifying the AST. Neither the source code is modified nor
new Java files are generated. Moreover, and probably more importantly,
Lombok supports generation of code \emph{inside} a class/interface,
which conventional Java annotation processors do not support. For
example, the standard \texttt{javax.annotation} processor, which is part of the
Java platform, only allows generation of \emph{new code}, and the
new code has to be written in \emph{new files}. Modification and/or
reinterpretation of existing code is not supported.

@@

\begin{figure*}[t]
\saveSpaceFig
\saveSpaceFig
\centering
\includegraphics[width=6in]{pdfs/screenshot4.png}
\caption{Generated methods shown in the Outline window of Eclipse and auto-completion.}
\label{fig:screenshot}
\saveSpaceFig
\end{figure*}

@@

\paragraph{Clarity against obfuscation}
Preprocessors bring great power, which can easily be misused producing
code particularly hard to understand. Thus code quality and maintainability are reduced.
Compilation agents start from Java syntax, but they can reinterpret it.
Preserving the syntax avoids syntactic conflicts, and allows many
tools to work transparently.

@@

\paragraph{Syntax and type errors}
Some preprocessors (like the C one) can produce syntactically invalid code.
Lombok ensures only syntactically valid code is produced. %; however, type errors can appear.
Classless Java additionally guarantees that no type errors are introduced
in generated code and client code. We discuss these two guarantees in
more detail next:

@@

\item{\bf Self coherence}: the generated code itself is well-typed. That is,
  type errors are not present in code the user has not written (for
  example \texttt{of} methods in Figure~\ref{fig:screenshot}).
In our case it means that either \mixin{} produces (in a controlled way) an
understandable error, or the interface can be successfully annotated and the generated code is well-typed.

\item{\bf Client coherence}: all the client code (for example method calls)
  that is well-typed before code generation is also well-typed after the generation.
The annotation just adds more behaviour without removing any functionality.

@@

\paragraph{Heir coherence} Another form of guarantee that could be
useful in AST rewriting is heir coherence. That is, interfaces
(and in general classes) inheriting the instrumented code are
well-typed if they were well-typed without the instrumentation.
In a strict sense, our rewriting \emph{does not} guarantee heir coherence.  The reason
is that this would forbid adding any (default or abstract) method to
the annotated interfaces, or even doing type refinement. Indeed consider
the following:

### 2016-03-24 OOPSLA16/sections/Introduction.tex
unify title/paragraph/enumerate capitalization: The rule is: - Capitalization All Section/Subsection Titles. - No capitalization for paragraph/enumerate titles. - No capitalization for the first word after a colon ":" (There's difference in this point between American and British English, I think we should stick to one).

\item {\bf Classless Java:} A practical realization of IB in
  Java. Classless Java is implemented using annotation processing, 
  allowing most tools to work transparently with our approach. 
  We provide examples, informal and formal descriptions of Classless
  Java.

\item {\bf Type-Safe Covariant Mutable State:} We show how the 
 combination of abstract state operations and type-refinement enables 
 a form of mutable fields that can be covariantly refined in a type-safe way.

@@

\item {\bf Case Studies:} We illustrate the usefulness of IB
  through various examples and case studies.

>>>>

\item {\bf Classless Java:} a practical realization of IB in
  Java. Classless Java is implemented using annotation processing, 
  allowing most tools to work transparently with our approach. 
  We provide examples, informal and formal descriptions of Classless
  Java.

\item {\bf Type-safe covariant mutable state:} we show how the 
 combination of abstract state operations and type-refinement enables 
 a form of mutable fields that can be covariantly refined in a type-safe way.

@@

\item {\bf Case studies:} we illustrate the usefulness of IB
  through various examples and case studies.

### 2016-03-24 OOPSLA16/sections/RelatedWork.tex
unify title/paragraph/enumerate capitalization: The rule is: - Capitalization All Section/Subsection Titles. - No capitalization for paragraph/enumerate titles. - No capitalization for the first word after a colon ":" (There's difference in this point between American and British English, I think we should stick to one).

\paragraph{Models of Object-Oriented Programming}
There are traditionally two big styles of OOP: class-based and
prototype-based. The class-based tradition started with Simula~\cite{Dahl66simula}, 
and is the dominant style of OOP today. Major languages such 
as Java, C\#, C++ or Scala favour a class based OOP style. Very often
class-based languages are statically typed and they have interfaces or
some other similar mechanisms such as traits. However, in contrast to IB
programming, interfaces or traits in CB programming are not meant to
replace classes, but rather, they complement classes. Normally
interfaces or traits define the signatures of the methods and
\emph{may} have some behaviour (such as in Java8 interfaces or Scala
traits). However classes are still responsible for object
instantiation and handling state. Prototype-based languages started
with Self~\cite{Ungar87self}, and are still widely used today via languages like
JavaScript. Frequentely prototype-based languages are dynamically 
typed. IB programming is clearly closer to CB programming. However the
key difference is that classes are not needed in IB
programming. Object interfaces recover the responsability of object
instantiation, and abstract state operations provide a way to deal
with state. A primary motivation of IB programming is to improve 
modularity and flexibility. Central to these goals are the 
more expressive and automated forms of type-refinement, and support 
for multiple interface inheritance.

\paragraph{Multiple Inheritance in Object Oriented Languages}
Many authors have argued in favor or against multiple inheritance.
Multiple \emph{implementation}
inheritance is very expressive, but difficult to model and implement, and can
cause difficulties (including the famous diamond (fork-join)
problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.) in reasoning about
programs. To conciliate the need for expressive power and simplicity, many
models have been proposed, including C++ virtual inheritance,
mixins~\cite{bracha90mixin}, traits~\cite{scharli03traits}, and hybrid models
such as CZ~\cite{malayeri2009cz}.  They provide novel programming architecture
models in the OO paradigm. In terms of restrictions set on these models, C++
virtual inheritance aims at a relative general model; the mixin model adds some
restrictions; and the trait model is the most restricted one (excluding
state, instantiation, etc).

>>>>

\paragraph{Models of Object-Oriented programming}
There are traditionally two big styles of OOP: class-based and
prototype-based. The class-based tradition started with Simula~\cite{Dahl66simula}, 
and is the dominant style of OOP today. Major languages such 
as Java, C\#, C++ or Scala favour a class based OOP style. Very often
class-based languages are statically typed and they have interfaces or
some other similar mechanisms such as traits. However, in contrast to IB
programming, interfaces or traits in CB programming are not meant to
replace classes, but rather, they complement classes. Normally
interfaces or traits define the signatures of the methods and
\emph{may} have some behaviour (such as in Java8 interfaces or Scala
traits). However classes are still responsible for object
instantiation and handling state. Prototype-based languages started
with Self~\cite{Ungar87self}, and are still widely used today via languages like
JavaScript. Frequentely prototype-based languages are dynamically 
typed. IB programming is clearly closer to CB programming. However the
key difference is that classes are not needed in IB
programming. Object interfaces recover the responsability of object
instantiation, and abstract state operations provide a way to deal
with state. A primary motivation of IB programming is to improve 
modularity and flexibility. Central to these goals are the 
more expressive and automated forms of type-refinement, and support 
for multiple interface inheritance.

\paragraph{Multiple inheritance in Object-Oriented languages}
Many authors have argued in favor or against multiple inheritance.
Multiple \emph{implementation}
inheritance is very expressive, but difficult to model and implement, and can
cause difficulties (including the famous diamond (fork-join)
problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.) in reasoning about
programs. To conciliate the need for expressive power and simplicity, many
models have been proposed, including C++ virtual inheritance,
mixins~\cite{bracha90mixin}, traits~\cite{scharli03traits}, and hybrid models
such as CZ~\cite{malayeri2009cz}.  They provide novel programming architecture
models in the OO paradigm. In terms of restrictions set on these models, C++
virtual inheritance aims at a relative general model; the mixin model adds some
restrictions; and the trait model is the most restricted one (excluding
state, instantiation, etc).

### 2016-03-24 OOPSLA16/sections/Translation.tex
unify title/paragraph/enumerate capitalization: The rule is: - Capitalization All Section/Subsection Titles. - No capitalization for paragraph/enumerate titles. - No capitalization for the first word after a colon ":" (There's difference in this point between American and British English, I think we should stick to one).

\paragraph{Type preservation:}
Note that we preferred to introduce self, client and heir coherence instead of referring to conventional
type preservation theorems. The reason is
to better model how our approach behaves in a object oriented software ecosystem with inheritance,
where only some units may be translated/expanded.
Note inheritance's crucial influence in heir coherence.
Our formulation of client coherence allows us to discuss about intermediate
stages where only some code units are translated/expanded. Conventional 
type preservation refers only to completely translated program.
Our coherence guarantees mean that developers and designers of Java libraries and frameworks
can start using IB (and our \mixin annotation) in the evolution of their products
and still retain backward compatibility with their clients.

>>>>

\paragraph{Type preservation}
Note that we preferred to introduce self, client and heir coherence instead of referring to conventional
type preservation theorems. The reason is
to better model how our approach behaves in a object oriented software ecosystem with inheritance,
where only some units may be translated/expanded.
Note inheritance's crucial influence in heir coherence.
Our formulation of client coherence allows us to discuss about intermediate
stages where only some code units are translated/expanded. Conventional 
type preservation refers only to completely translated program.
Our coherence guarantees mean that developers and designers of Java libraries and frameworks
can start using IB (and our \mixin annotation) in the evolution of their products
and still retain backward compatibility with their clients.
